### 2015-06-10 ncdnpaper/LinearPrograms2.tex
Initial import of thesis starting from proposal document
Includes all files: tex, pdf, etc.

\section{NCDN management strategies}
\label{sec:optimize}

%\textcolor{blue}{
%Content-aware traffic engineering for an NCDN refers to the combined task of content distribution and traffic engineering. There are several content-aware traffic engineering strategies that an NCDN can deploy. An NCDN can use any combination of demand-aware or demand-oblivious content distribution and routing strategies (Section \ref{sec:background})  as its content-aware traffic engineering strategy. An NCDN can also jointly optimize both content distribution and traffic engineering in a demand-aware manner.
%}

To answer the above questions, we develop an optimization model for NCDNs to jointly optimize placement, routing, and redirection so as to optimize network cost or user-perceived latency. We formulate the optimization problem as a mixed-integer program (MIP), present hardness and inapproximability results, and discuss approximation heuristics to solve MIPs for realistic problem sizes.

%\textcolor{blue}{
%This section presents a optimization strategy for NCDNs to jointly optimize content placement, routing, and redirection. First we describe our NCDN model and cost functions followed by a description of the optimization formulation as a mixed-integer program (MIP). Next, we present hardness and inapproximability results for this problem. Finally, we discuss approximation techniques that we use to solve our MIP formulation. 
%}

\eat{
In this section, we formalize the \ncp\ model and the problem of determining the optimal content placement and routing as a mixed integer program (MIP).  
Next, we present hardness and inapproximability results for this problem. Finally, we discuss approximation techniques that we use to solve our MIP formulation. 
}

%This formulation takes as input a content matrix, i.e., the demand for each content at each network point-of-presence (PoP), and  computes a placement and a routing that minimizes the maximum link utilization (MLU) while respecting link capacity and storage constraints.

%Further, we present a variant of this MIP that determines the optimal placement strategy for a given routing scheme. 
%We also present a variant of the MIP that determines the optimal placement strategy for a given routing scheme

@@

%In this section, we formalize the \ncp\ model and present three strategies to enable placement-routing-redirection in \ancp.  First, we present a mixed integer program (MIP) which calculates  content placement and routing to optimize network cost based a content matrix, i.e., the demand for each piece of content at each network point-of-presence (PoP). 
%Next, we explain why content chunking, i.e., delivering content in small chunks, can improve performance of  both demand aware and demand oblivious content placement. Finally, we describe a link load aware request redirection strategy for a demand oblivious placement, designed to reduce \ancp's network cost.

@@

\begin{table}[t]
\begin{small}
\begin{center}
\begin{tabular}{|p{4.65in}|} 
 \hline
\textbf{Input variables and descriptions}\\ 
\end{tabular}
\begin{tabular}{|p{0.2in}|p{4.3in}|}
\hline
  $V$ & Set of nodes where each node represents a PoP   \\
  $E$ & Set of edges where each link represents a communication link   \\
   $o$ & Virtual origin node that hosts all the content in $K$  \\ 
%  $ X(o)$ & $X(o) \in V$  s.t., $o \in O$ is connected to $X(o)$ with infinite capacity link.\\ 
   % $ X$ & $X  \subset V$.   Each $x\in X$ is connected to unique $o \in O$  and vice versa.\\ 
  $ X$ & Set of exit nodes in $V$ \\ 
  $D_i$ & Disk capacity at node $i \in V$ (in bytes) \\
  $C_e$ & Capacity of link $e \in E$ (in bits/sec)\\
  $K$ & the set of all content accessed by end-users \\
   $S_k$ & Size of content $k \in K$.   \\
   $T_{ik}$ & Demand (in bits/sec)  at  node $i \in V$ for content $k \in K$ \\
  % $P(i)$ & Set of  outgoing links at PoP $i \in V$\\
   %$Q(i)$ & Set of  incoming links at PoP $i \in V$\\
   \hline
   \end{tabular}
\begin{tabular}{|p{4.65in}|}
  \hline
\textbf{Decision variables and descriptions}\\ 
\end{tabular}
\begin{tabular}{|p{0.2in}|p{4.3in}|}
\hline
$\alpha$ & MLU of the network \\
$z_k$ & Binary variable indicating whether one or more copies of content $k$ is placed in the network\\
$x_{jk}$ & Binary variable indicating whether content $k$  is placed at node $j \in V \cup \{o\}$\\
$f_{ij}$ & Total traffic from node $j$ to node $i$\\
$f_{ije}$ & Traffic from node $j$ to node $i$ crossing link $e$.\\
$t_{ijk}$ & Traffic demand at node $i \in V$ for content $k\in K$ served from node $j \in V \cup \{o\}$ \\\hline
\end{tabular}
\end{center}
\end{small}
\label{table:paramtable}
\caption{List of input  and decision variables for the \ncp\ problem formulation.}
\vspace{-0.2in}
\end{table}

\subsection{\ncp\ model}
\label{sec:model}

@@

Table 1 lists all the model parameters. \Ancp\ consists of a set of nodes $V$ where each node represents a PoP in the network. The nodes are connected by a set of directed edges $E$ that represent the backbone links in the network. The set of content requested by end-users is represented by the set $K$ and the sizes of content are denoted by $S_k, k\in K$.  The primary resource constraints are the link capacities $C_e, e\in E$, and the storage at the nodes $D_i, i\in V$. We implicitly assume that the content servers at the PoPs have adequate compute resources to serve locally stored content.

@@

A {\em content matrix} (CM) specifies the demand for each content at each node. An entry in this matrix, $T_{ik}, i\in V, k\in K$, denotes the demand (in bits/second) for content $k$ at node $i$.  CM is assumed to be measured by the \ncp\ a priori over a coarse-grained interval, e.g., the previous day.  The infrastructure required for this measurement is comparable to what ISPs have in place to monitor traffic matrices today.

Origin servers, owned and maintained by the NCDN's content providers, initially store all content published by content providers. We model origin servers using a single virtual origin node $o$ external to the \ncp\ that can be reached via a set of exit nodes $X \subset V$ in the \ncp\ (Figure~\ref{fig:NCDNArch}). 
Since we are not concerned with traffic engineering links outside the \ncp, we model the edges $(x, o)$, for all $x \in X$, as having infinite capacity. The virtual origin node $o$  always maintains a copy of all the requested content. However, a request for a content is served from the virtual origin node only if no copy of the content is stored at any node $i \in V$. In this case, the request is assumed to be routed to the virtual origin  via the exit node closest to the node where the request was made (in keeping with the commonly practiced {\em early-exit} or {\em hot potato} routing policy).

%\textcolor{blue}
%{
ISP networks carry transit traffic in addition to NCDN traffic, which can be represented as a transit traffic matrix (TTM).  Each entry in the TTM contains the volume of transit traffic between two PoPs in the network.

%ISPs today optimize routing for TTM using traditional TE methods. 
%A simple extension to the joint optimization described next can jointly optimize routing for TTM and CM. We omit the extension for ease of exposition.  Further, our findings presented in Section \ref{sec:eval} suggest that a joint optimization of CM and TTM is not necessary for NCDNs.
%}

%All content published by content providers of the \ncp\ service is initially stored at origin servers owned and maintained by the content providers. We model origin servers using a single virtual origin node $o$ external to the \ncp\ that can be reached from any node $i \in V$ by routing to the closest exit node $x \in X$, where $X \subset V$ is the set of all exit nodes in the \ncp\ (see Figure~\ref{fig:NCDNArch}). Since we are not concerned with traffic engineering links outside the \ncp, we model the edges $(x, o)$, for all $x \in X$, as having infinite capacity. The virtual origin node $o$  always maintains a copy of all the requested content. However, a request for a content is served from the virtual origin node only if no copy of the content is stored at any node $i \in V$. In this case, the request is assumed to be routed to the virtual origin  via the exit node closest to the node where the request was made (in keeping with the commonly practiced {\em early-exit} or {\em hot potato} routing policy).

%All content is published by content providers of the \ncp\ service and initially stored at a set of origin servers owned and maintained by the content providers. We assume that origin servers are hosted external to the \ncp. These origin servers are typically mirrored across different data centers and multihomed across different networks.  For simplicity, we model a single virtual origin node $o$ external to the \ncp\ that can be reached from any node $i \in V$ by routing to the closest exit node $x \in X$, where $X \subset V$ is the set of all exit nodes in the \ncp\ (See Figure~\ref{fig:NCDNArch}). Since we are not concerned with traffic engineering links outside the \ncp, we model the edges $(x, o)$, for all $x \in X$, as having infinite capacity. The virtual origin node $o$  always maintains a copy of all the requested content. Nodes in the network may additionally store local copies of some of the content. A request for a content is served from the virtual origin node only if no copy of the content is stored at any node in $V$ of the \ncp. In this case, the request is assumed to be routed to the virtual origin  via the exit node closest to the node where the request was made (in keeping with the commonly practiced {\em early-exit} or {\em hot potato} routing policy).

%The customers of the CDN service, e.g. media companies, publish content which are initially stored at origin servers ($O$)  owned by the customer which are located outside the ISP network. Every origin server has a copy of all the content published. But the origin server is contacted only if no copy of the content is available in the network. Traffic going to/from an origin server exits/enters the network at a unique node. The set of nodes which serve as the entry/exit point of traffic between origin servers and network is denoted by set $X$. A user's request is sent to the origin node with the least hop count distance  from the node at which the user is present; ties are broken randomly.

%The set of content being served is denoted by the set $K$. Each content $k$ from this set has a finite size $S_k$ and a measurable demand $T_{ik}$ at each node $i \in V$. Demand for a content at a node is measured in bits/second and is calculated by dividing the aggregate traffic  measured in bits in a given interval with the length of the interval in seconds.

%Our optimization objective is to minimize the maximum load on all links and the maximum load on all network nodes.

%Our network model resembles an architecture likely to be adopted by a Network CDN.  The network consists of a set of point of presence locations (PoPs), which are interconnected by high capacity links which form the backbone of the network.  To provide CDN services through its network, the network has storage and servers co-located at all the PoPs in the network.

%Since a Network CDN continues to provide IP transit services, we sometimes use the term ISP in place of Telco CDN.

%The customers of the CDN service, e.g. media companies, publish content which are initially stored at an origin server owned by the customer which may be located outside the ISP network. Depending on content popularity and locations where it is popular, the content is pulled from origin servers  and stored at one or more PoPs. The users request for the content is first received by the nearest PoP and is either served locally or is redirected to another PoP which has a copy of the content. A request is send to the origin server if the content isn't available in the network. The content stored at each PoP can change over time depending on  a change in  content popularity patterns  and the strategy used by Network CDN.

%\noindent\textbf{Network-CDN controls content placement}

%A Network-CDNÊdiffers from a non-CDN ISP in that it controls both the \emph{content placement} and routing in the network while a non-CDN ISP only controls routing. By content placement, we mean the strategy which decides  where is each content to placed, and where should each users' request for a content be redirected.

%\noindent\textbf{imp. assumption: only consider CDN traffic, no transit traffic}

%A Network-CDN will generate intradomain as well as interdomain traffic. We focus only on the intradomain traffic generated by the CDN service. We assume that  intradomain IP transit traffic is handled independently. 
% \footnote[1]{Such a mechanism can be implemented by the ISP by isolating CDN and IP transit traffic using MPLS \cite{cisco-tagswitching}. An ISP can isolate the CDN traffic and IP transit traffic using MPLS and adopt independent traffic engineering schemes for each of them.}

%
%The objective of the TE problem is to calculate a content placement and routing so as to optimize a network cost function, which in our case is the maximum link utilization or MLU.

%Network and content provider  or \ncp\  stores content at each PoP in the network and has to pay for upgrading its link capacity. The content distributor is a profit-maximizing entity. Assuming that the cost of upgrades is the dominant cost of operating the network, its objective is to maximize the amount of traffic it can route with the current infrastructure. We assume a linear scaling of traffic demands, and thus we seek to maximize the increase in traffic demands which can be supported with the current infrastructure.

%Given a network infrastructure, an ISPs problem consists of three sub-problems: (1) Where to place content given storage constraints ? (2) How to split traffic demands among multiple copies of content available at different locations? (3) How to compute routing in the network ?

%Our network model consists of following parameters and variables:

%\noindent\textbf{Parameters:}
%\begin{itemize}
%\item
%Network  $G = (V,E)$.
%\item
%Nodes $V = \{ 1, 2, ... \}$
%\item
%Links $ E = \{ e_{ij}\}$ for link between node $i$ and  $j$ and link capacity = $ C = \{ c_{ij}\}$.
%\item
%Disk space at node $i$, $D_i$. 
%\item
%Size of content, $S = \{ s_1, s_2, ...\}$, $s_j$ is the size of content $j$.
%\item
%Aggregate demand for content is $F = \{f_1, f_2, ... \}$ , $f_j$ is the total demand for content $j$.
%\item
%%Traffic demands $F = \{f_1, f_2, ... \}$ , $f_i$ is the total demand at node $i$.
%Demand at each node for content $j$, i.e., $f_j = f_{1j} + f_{2j} + ... f_{ij}$ , $f_{ij}$ is the total demand for content $j$ at node $i$.
%\end{itemize}
%
%There are three metrics which an \ncp\ needs to consider while calculating placement and routing. (1) Link utilization: Traditional traffic engineering 
%
%1. load at servers
%2. delay of users
%3. link utilization
%
%
%We assume that storage available at different infrastructures is given. 
%
%%Destination for flow $f_i$ is node  $t_i$.
%
%\noindent\textbf{Variables:}
%\begin{itemize}
%\item
%$\alpha$, the maximum link utilization.
%\item
%$x_{jk}$ - If $x_{jk} = 1$ node $k$ is a hosting content $j$ and $x_{jk} = 0$ otherwise.
%\item
%$y_{ik}$ - Aggregate traffic from source $k$ to sink $i$ for all content in the network.
%\item
%$y_{ijk}$ - Traffic from source $k$ to sink $i$ for content $j$.
%\item
%$h_{ik}^{pq}$ - Traffic on link $pq$ from source $k$ to sink $i$ .
%\end{itemize}
%
%In some of our traffic engineering schemes, some variables are known parameters.
%e.g., if the placement is fixed then $y_{ijk}$  and $x_{jk}$ are known
%parameters instead of variables.
%
%\noindent\textbf{Optimization objective}
%
%We solve the following problem: given a network and storage infrastructure, we define following terms which capture the load on the network.
%
%(1) Load on a link: This is the ratio of traffic on a link to the capacity of the link. 
%
%(2) Load on a network node: (Current load)/(Maximum supported load)
%
%
%
%There are three resources in the network: (1) capacity of links between nodes. (2) storage  at each node and (3) maximum supportable workload of web servers at each node. The units for link capacity (Gbps, Mbps) and storage at each node (GB, MB) are straightforward. As a simplistic assumption, we assume the workload at each node is proportional to the outgoing traffic from servers at each node. The capacity of web servers is written in terms of the maximum data rate which such a network can support. 
%
%While an \ncp\ would want to use all the available  storage at each node, the same is not true of the other two resources. ISP's prefer running network links at low utilization levels and same would hold true for \ncp's. Web servers of content distribution networks, are also kept running at low utilization.

% files stored
% # requests stored

%We model the load on network links and servers at each node as follows: (1) The load on a link in the network is equal to the utilization of the link, i.e., the traffic on  a link divided by the capacity of the link.

%Load = (current traffic)/(maximum supportable traffic)
%
%ISP's use this definition of load in traffic engineering. (2) The load on servers at each node is assumed to be proportional to the traffic served from that node.

%The load on a link may not be vis-a-vis comparable with the load on a network node. Therefore, we multiply the load on a network node with a normalizing constant.

\subsection{Cost functions}
\label{sec:costfunction}

We evaluate NCDN-management strategies based on two cost functions. The
first cost function is maximum link utilization (or MLU) which measures the effectiveness of traffic engineering in an NCDN. MLU is a widely used network cost function for traditional TE.

%the sum of latency of transfers in the network, which is equivalent
The second cost function models user-perceived latency and is defined as  $\sum_{e \in E} X_e$, where $X_e$ is the product of traffic on link $e$ and its link latency $L(e)$.  The latency of a link $L(e)$ is the sum of  a fixed propagation delay and a variable utilization dependent delay. For a unit flow, link latency is defined as $ L_e(u_e) = p_e (1 + f(u_e))$, where $p_e$ is the propagation delay of edge $e$, $u_e$ is its link utilization, and $f(u)$ is a piecewise-linear convex function.
%
%% defined as $f(0) = 0$ and its first derivatives:
%
%\begin{eqnarray}
%  f'(u)  =  \begin{cases}  0 &  u \leq 0.6,  \notag \\
%  2 & 0.6 < u \leq 0.7,  \notag \\
%    4 & 0.7 < u \leq 0.8,  \notag \\
%      16 & 0.8 < u \leq 0.9,  \notag \\
%      128 & 0.9 < u \leq 1.0,  \notag \\      
%  2048 &  1  < u, \end{cases} \end{eqnarray}
%  
This cost function is similar to that used by Fortz and Thorup \cite{fortz2000internet}. At small link utilizations ( $< 0.6$), link latency is determined largely by propagation delay hence $f$ is zero. At higher link utilizations (0.9 and above) an increase in queuing delay and delay caused by retransmissions significantly increase the effective link latency. The utilization-dependent delay is modeled as proportional to propagation delay as the impact of (TCP-like) retransmissions is more on paths with longer links.  Since $L_e$ is convex, a set of linear constraints can be written to constraint the value of $X_e$  (as in \cite{fortz2000internet}). %We omit these constraints here due to lack of space.

%The latency cost function is  $\sum_{e \in E} X_e$ where $X_e$ is the product of traffic on link $e$ and its link latency, $L(e)$.

\subsection{Optimal strategy as MIP}
\label{sec:linearprograms}

%The problem of placment-routing-redirection by \ncp s ({\em \ncp\ problem} for short) seeks to compute a placement and routing that minimizes the MLU and satisfies the demands specified by  the content matrix while respecting link capacity and storage constraints. This optimization goal can be formulated as a mixed integer program (MIP). Unlike the traditional traffic engineering problem that can be formulated as a multi-commodity flow problem and solved using a linear program, the \ncp\ problem needs to make binary placement decisions, i.e., whether or not to place a content at a PoP, and then route the demand accordingly.

We present here a joint optimization strategy for NCDN-management formulated as a MIP. 
This formulation takes as input a content matrix, i.e., the demand for each content at each network point-of-presence (PoP), and  computes content placement, request redirection and routing that minimizes an NCDN cost function  while respecting link capacity and storage constraints.
The decision variables for this problem are listed in  Table 1. The MIP to minimize an NCDN cost function $C$ (either MLU or latency) is as follows:
\begin{equation}
\min  C
\end{equation}
\begin{equation*}
\mbox{subject to}
\end{equation*}
\begin{eqnarray}
\sum_{j \in V }  t_{ijk} +   t_{iok} &=& T_{ik}, \quad \forall k \in K, i \in V\\
\sum_{k \in K} t_{ijk} &=& f_{ij} ,  \ \forall j \in V-X, i \in V\\
\sum_{k \in K} t_{ijk} + \sum_{k \in K} \delta_{ij}  t_{iok} &=& f_{ij} ,  \ \forall j \in X, i \in V
\end{eqnarray}
where $\delta_{ij}$ is 1 if $j$ is the closest exit node to $i$ and 0 otherwise. Note that $\delta_{ij}$ is not a variable but a constant that is determined by the topology of the network, and hence constraint (4) is linear.
\begin{eqnarray}
 \sum_{p \in P(l)} f_{ijp} - \sum_{q \in Q(l)}  f_{ijq}  =  \begin{cases}  f_{ij} & \text{if } l=i,  \notag \\
   -f_{ij}  &   \text{if } l=j,  \notag \\
0 & \text{otherwise}, \end{cases}\\
\forall i,j,l \in V
\end{eqnarray}
where $P(l)$ and $Q(l)$ respectively denote the set of outgoing and incoming links at node $l$.
\begin{eqnarray}
 \sum_{i \in V, j \in V} f_{ije} &\leq& \alpha \times C_e, \quad \forall e \in E\\
 \sum_{k \in K}  x_{ik}S_k &\leq& D_i , \quad \forall i \in V\\
x_{ok} &=& 1, \quad \forall  k \in K \\
\sum_{i \in V}  x_{ik} &\geq& z_k, \quad \forall k \in K 
\end{eqnarray}
 \begin{eqnarray}
x_{ik} &\leq& z_k, \quad \forall k \in K, i \in V  
\end{eqnarray}
\vspace{-0.25in}
\begin{eqnarray}
t_{ijk} &\leq&  x_{jk} T_{ik},  \quad \forall k \in K,  i \in V, j \in V \cup \{o\}\\
t_{iok} &\leq& T_{ik}(1 - z_k), \quad  \forall   k \in K 
\end{eqnarray}
\vspace{-0.25in}
\begin{eqnarray*}
x_{jk}, \ z_k &\in& \{0,1\}, \quad \forall j \in V, k \in K\\
f_{ije}, \ t_{ijk}, \ t_{iok} &\geq& 0, \quad \forall  i,j \in V, e \in E, k\in K
\end{eqnarray*}

@@

The constraints have the following rationale.
Constraint (2) specifies that the total traffic demand at each node for each content must be satisfied.
Constraints (3) and (4) specify that the total traffic from source $j$ to sink $i$ is the sum over all content $k$ of the traffic from $j$ to $i$ for $k$.
Constraint (5) specifies that the volume of a flow coming in must equal that going out at each node other than the source or the sink.
Constraint (6) specifies that the total flow on a link is at most $\alpha$ times capacity.
Constraint (7) specifies that the total size of all content stored at a node must be less than its disk capacity.
Constraint (8) specifies that all content is placed at the virtual origin node $o$.
Constraints (9) and (10) specify that at least one copy of content $k$ is placed within the network if $z_k = 1$, otherwise $z_k = 0$ and no copies of $k$ are placed at any node. Constraint (11) specifies that the flow from a source to a sink for some content should be zero if the content is not placed at the source (i.e., when $x_{jk} = 0$), and the flow should be at most the demand if the content is placed at the source  (i.e., when $x_{jk} = 1$).
Constraint (12) specifies that if some content is placed within the network, the traffic from the origin for that content must be zero.

Updating the content placement itself generates traffic and impacts the link utilization in the network. For ease of exposition, we have deferred a formal description of the corresponding constraints to a techreport  \cite{techreportNCDN}.  Finally, a simple extension to this MIP presented in a techreport \cite{techreportNCDN}  jointly optimizes routing given a TTM as well a CM.  We have presented a CM-only formulation here as our findings (in Section \ref{sec:eval}) show that a joint optimization of the CM and TTM is not useful for NCDNs.
%\subsubsection{ISP Transit Traffic Optimization}

\subsection{Computational hardness}
\optloc\ is the decision version of the NCDN problem. The proofs for these theorems are presented in a techreport  \cite{techreportNCDN}.
%We show that the NCDN-problem is NP-Complete and is inapproximable to within an exponential factor.

{{\sc Theorem 1.}} {\em \optloc\ is NP-Complete even in the special case where all objects have unit size, and all demands, link capacities, and storage capacities have binary values.}

@@

{{\sc Corollary 1.}} {\em  \optloc\ is inapproximable to within a constant factor  unless P = NP.}

@@

\subsection{Approximation techniques for MIP}
\label{sec:approx}

@@

The first is a two-step local search technique. In the first step, we ``relax'' the MIP  by allowing the integral variables $x_{jk}$ and $z_k$ to take fractional values between $0$ and $1$. This converts an MIP into an LP that is more easily solvable. Note also that the optimal solution of the relaxed LP is a lower bound on the optimal solution of the MIP.  However, the LP solution may contain fractional placement of some of the content with the corresponding $x_{jk}$ variables set to fractional values between $0$ and $1$.  However, in our experiments only about $20\%$ of the variables in the optimal LP solution were set to fractional values between $0$ or $1$, and the rest took integral values of $0$ or $1$.  In the second step, we search for a valid solution for the MIP in the local vicinity of the LP solution by substituting the values for variables that were set to $0$ or $1$ in the LP solution, and re-solving the MIP for the remaining variables. Since the number of integer variables in the second MIP is much smaller, it can be solved more efficiently than the original MIP.

@@

A \unplanned\ placement improves with content chunking due to two reasons: (1) A chunk is downloaded much before the complete file is downloaded, and hence it can be uploaded to other caches sooner. (2) Content chunking enables caches to store a partially downloaded content if the user aborts the download before completion. Without content chunking support, partially downloaded content is useless and hence must be discarded from cache. The alternative is to download the remaining fraction of content to the cache even though the user has aborted the download. Both these choices lead to a wastage of network bandwidth and the later also cause unnecessary cache evictions.  Splitting large content into smaller chunks significantly reduces network bandwidth wastage, reduces avoidable cache evictions and improves network cost.

>>>>

\eat
{
\section{Content-Aware Traffic\\ Engineering for An NCDN}
\label{sec:optimize}

In this section, we formalize the \ncp\ model and the problem of determining the optimal content placement and routing as a mixed integer program (MIP).  This formulation takes as input a content matrix, i.e., the demand for each content at each network point-of-presence (PoP), and  computes a placement and a routing that minimizes the maximum link utilization (MLU) while respecting link capacity and storage constraints.
}
%Further, we present a variant of this MIP that determines the optimal placement strategy for a given routing scheme. 
%We also present a variant of the MIP that determines the optimal placement strategy for a given routing scheme

@@

%In this section, we formalize the \ncp\ model and present three strategies to enable content-aware traffic engineering in \ancp.  First, we present a mixed integer program (MIP) which calculates  content placement and routing to optimize network cost based a content matrix, i.e., the demand for each piece of content at each network point-of-presence (PoP). 
%Next, we explain why content chunking, i.e., delivering content in small chunks, can improve performance of  both demand aware and demand oblivious content placement. Finally, we describe a link load aware request redirection strategy for a demand oblivious placement, designed to reduce \ancp's network cost.

@@

\begin{table}[t]
\begin{small}
%\begin{center}
\begin{tabular}{|p{3.15in}|} 
 \hline
\textbf{Input variables and descriptions}\\ 
\end{tabular}
\begin{tabular}{|p{0.2in}|p{2.8in}|}
\hline
  $V$ & Set of nodes where each node represents a PoP   \\
  $E$ & Set of edges where each link represents a communication link   \\
   $o$ & Virtual origin node that hosts all the content in $K$  \\ 
%  $ X(o)$ & $X(o) \in V$  s.t., $o \in O$ is connected to $X(o)$ with infinite capacity link.\\ 
   % $ X$ & $X  \subset V$.   Each $x\in X$ is connected to unique $o \in O$  and vice versa.\\ 
  $ X$ & Set of exit nodes in $V$ \\ 
  $D_i$ & Disk capacity at node $i \in V$ (in bytes) \\
  $C_e$ & Capacity of link $e \in E$ (in bits/sec)\\
  $K$ & the set of all content accessed by end-users \\
   $S_k$ & Size of content $k \in K$.   \\
   $T_{ik}$ & Demand (in bits/sec)  at  node $i \in V$ for content $k \in K$ \\
  % $P(i)$ & Set of  outgoing links at PoP $i \in V$\\
   %$Q(i)$ & Set of  incoming links at PoP $i \in V$\\
   \hline
   \end{tabular}
\begin{tabular}{|p{3.15in}|}
  \hline
\textbf{Decision variables and descriptions}\\ 
\end{tabular}
\begin{tabular}{|p{0.2in}|p{2.8in}|}
\hline
$\alpha$ & MLU of the network \\
$z_k$ & Binary variable indicating whether one or more copies of content $k$ is placed in the network\\
$x_{jk}$ & Binary variable indicating whether content $k$  is placed at node $j \in V \cup \{o\}$\\
$f_{ij}$ & Total traffic from node $j$ to node $i$\\
$f_{ije}$ & Traffic from node $j$ to node $i$ crossing link $e$.\\
$t_{ijk}$ & Traffic demand at node $i \in V$ for content $k\in K$ served from node $j \in V \cup \{o\}$ \\\hline
\end{tabular}
\label{table:paramtable}
\end{small}

\caption{List of input  and decision variables for the \ncp\ problem formulation.}
\vspace{-0.2in}
\end{table}

\section{\ncp\ Model}
\label{sec:model}

@@

We formally describe our NCDN model in this section. Table \ref{table:paramtable} lists all the model parameters. \Ancp\ consists of a set of nodes $V$ where each node represents a PoP in the network. The nodes are connected by a set of directed edges $E$ that represent the backbone links in the network. The set of content requested by end-users is represented by the set $K$ and the sizes of content are denoted by $S_k, k\in K$.  The primary resource constraints are the link capacities $C_e, e\in E$, and the storage at the nodes $D_i, i\in V$. We implicitly assume that the content servers at the PoPs have adequate compute resources to serve locally stored content.

@@

A {\em content matrix} specifies the demand for each content at each node. An entry in this matrix, $T_{ik}, i\in V, k\in K$, denotes the demand (in bits/second) for content $k$ at node $i$.  The content matrix is assumed to be measured by the \ncp\ a priori over a coarse-grained interval, e.g., the previous day. The infrastructure required for this measurement is comparable to what ISPs have in place for monitoring traffic matrices today.

All content published by NCDNs content providers is initially stored at origin servers owned and maintained by the content providers. We model origin servers using a single virtual origin node $o$ external to the \ncp\ that can be reached via a set of exit nodes $X \subset V$ in the \ncp\ (see Figure~\ref{fig:NCDNArch}). 
Since we are not concerned with traffic engineering links outside the \ncp, we model the edges $(x, o)$, for all $x \in X$, as having infinite capacity. The virtual origin node $o$  always maintains a copy of all the requested content. However, a request for a content is served from the virtual origin node only if no copy of the content is stored at any node $i \in V$. In this case, the request is assumed to be routed to the virtual origin  via the exit node closest to the node where the request was made (in keeping with the commonly practiced {\em early-exit} or {\em hot potato} routing policy).

\section{Content-Aware Traffic\\ Engineering for An NCDN}
\label{sec:optimize}

%\subsection{Optimal Strategy as a MIP}
\label{sec:linearprograms}

The content-aware traffic engineering problem for \ncp s (\ncp\ problem, for short) seeks to compute a placement and routing that minimizes the MLU and satisfies the demands specified by  the content matrix while respecting link capacity and storage constraints. This optimization goal can be formulated as a mixed integer program (MIP). Unlike the traditional traffic engineering problem that can be formulated as a multi-commodity flow problem and solved using a linear program, the \ncp\ problem needs to make binary placement decisions, i.e., whether or not to place a content at a PoP, and then route the demand accordingly. The placement decision variables and other decision variables  for this problem are listed in  Table \ref{table:paramtable}. The MIP to minimize the MLU $\alpha$ is as follows:
\begin{equation}
\min  \alpha
\end{equation}
subject to
\begin{eqnarray}
\sum_{j \in V }  t_{ijk} +   t_{iok} &=& T_{ik}, \quad \forall k \in K, i \in V\\
\sum_{k \in K} t_{ijk} &=& f_{ij} ,  \ \forall j \in V-X, i \in V\\
\sum_{k \in K} t_{ijk} + \sum_{k \in K} \delta_{ij}  t_{iok} &=& f_{ij} ,  \ \forall j \in X, i \in V
\end{eqnarray}
where $\delta_{ij}$ is 1 if $j$ is the closest exit node to $i$ and 0 otherwise. Note that $\delta_{ij}$ is not a variable but a constant that is determined by the topology of the network, and hence constraint (4) is linear.
\begin{eqnarray}
 \sum_{p \in P(l)} f_{ijp} - \sum_{q \in Q(l)}  f_{ijq}  =  \begin{cases}  f_{ij} & \text{if } l=i,  \notag \\
   -f_{ij}  &   \text{if } l=j,  \notag \\
0 & \text{otherwise}, \end{cases}\\
\forall i,j,l \in V
\end{eqnarray}
where $P(l)$ and $Q(l)$ respectively denote the set of outgoing and incoming links at node $l$.
\begin{eqnarray}
 \sum_{i \in V, j \in V} f_{ije} &\leq& \alpha \times C_e, \quad \forall e \in E \\
 \sum_{k \in K}  x_{ik}S_k &\leq& D_i , \quad \forall i \in V\\
x_{ok} &=& 1, \quad \forall  k \in K \\
\sum_{i \in V}  x_{ik} &\geq& z_k, \quad \forall k \in K  \\
x_{ik} &\leq& z_k, \quad \forall k \in K, i \in V  
\end{eqnarray}
\vspace{-0.25in}
\begin{eqnarray}
t_{ijk} &\leq&  x_{jk} T_{ik},  \quad \forall k \in K,  i \in V, j \in V \cup \{o\}\\
t_{iok} &\leq& T_{ik}(1 - z_k), \quad  \forall   k \in K 
\end{eqnarray}
\vspace{-0.25in}
\begin{eqnarray*}
x_{jk}, \ z_k &\in& \{0,1\}, \quad \forall j \in V, k \in K\\
f_{ije}, \ t_{ijk}, \ t_{iok} &\geq& 0, \quad \forall  i,j \in V, e \in E, k\in K
\end{eqnarray*}

@@

The constraints have the following rationale.
Constraint (2) specifies that the total traffic demand at each node for each content must be satisfied.
Constraints (3) and (4) specify that the total traffic from source $j$ to sink $i$ is the sum over all content $k$ of the traffic from $j$ to $i$ for $k$.
Constraint (5) specifies that the volume of a flow coming in must equal that going out at each node other than the source or the sink.
Constraint (6) specifies that the total flow on a link is at most $\alpha$ times capacity.
Constraint (7) specifies that the total size of all content stored at a node must be less than its disk capacity.
Constraint (8) specifies that all content is placed at the virtual origin node $o$.
Constraints (9) and (10) specify that at least one copy of content $k$ is placed within the network if $z_k = 1$, otherwise $z_k = 0$ and no copies of $k$ are placed at any node. Constraint (11) specifies that the flow from a source to a sink for some content should be zero if the content is not placed at the source (i.e., when $x_{jk} = 0$), and the flow should be at most the demand if the content is placed at the source  (i.e., when $x_{jk} = 1$).
Constraint (12) specifies that if some content is placed within the network, the traffic from the origin for that content must be zero. Updating the content placement itself generate traffic and impacts the MLU in the network. We have omitted a formal description of the corresponding constraints for ease of exposition and lack of space. %We describe our approach to limit the MLU while updating the content placement in the Appendix.

\subsection{Computational Hardness}
{{\sc Theorem 1}} {\em The NCDN-problem is NP-Complete even in the special case where all objects have unit size, and all demand are unit value.}

@@

{{\sc Theorem 2}} {\em  The NCDN problem is inapproximable within a $2^{O(I)}$ factor unless P = NP, where $I$ is the input size of the problem.}

The proofs for these theorems are included in \cite{techreport}.

@@

\subsection{Approximation Techniques}
\label{sec:approx}

@@

The first is a two-step local search technique. In the first step, we ``relax'' the MIP  by allowing the integral variables $x_{jk}$ and $z_k$ to take fractional values between $0$ and $1$. This converts a MIP into an LP that is more easily solvable. Note also that the optimal solution of the relaxed LP is a lower bound on the optimal solution of the MIP.  However, the LP solution may contain fractional placement of some of the content with the corresponding $x_{jk}$ variables set to fractional values between $0$ and $1$.  However, in our experiments only about $20\%$ of the variables in the optimal LP solution were set to fractional values between $0$ or $1$, and the rest took integral values of $0$ or $1$.  In the second step, we search for a valid solution for the MIP in the local vicinity of the LP solution by substituting the values for variables that were set to $0$ or $1$ in the LP solution, and re-solving the MIP for the remaining variables. Since the number of integer variables in the second MIP is much smaller, it can be solved more efficiently than the original MIP.

@@

A demand oblivious placement improves with content chunking due to two reasons: (1) A chunk is downloaded much before the complete file is downloaded, and hence it can be uploaded to other caches sooner. (2) Content chunking enables caches to store a partially downloaded content if the user aborts the download before completion. Without content chunking support, partially downloaded content is useless and hence must be discarded from cache. The alternative is to download the remaining fraction of content to the cache even though the user has aborted the download. Both these choices lead to a wastage of network bandwidth and the later also cause unnecessary cache evictions.  Splitting large content into smaller chunks significantly reduces network bandwidth wastage, reduces avoidable cache evictions and improves network cost.
